Abstractions for Multi-Sorted Substitutions

### 2024-03-13 paper/paper.tex
Update.

@@

  For this purpose, the \ACode{\ACon{CKit}} abstraction, needs
  to satisfy two requirements:
  it needs to provide us with enough machinery to define composition, but
  it also needs to provide us with enough axioms that we are able to instantiate
  the \ACode{\ACon{CTraversal}}. It is defined as followed:
  \LibCode\KCKit
  The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
  (according to \ACode{K₁}) and a renaming or substitution
  (according to \ACode{K₂}) and performs the map traversal returning
  a variable or term (according to \ACode{K₁⊔K₂}).
  The third parameter \ACode{K₁⊔K₂} can be seen as a functional
  dependency\cite{DBLP:conf/esop/Jones00} and is completely determined
  by the choice of \ACode{K₁} and \ACode{K₂}.

>>>>

  A compose kit \ACode{\ACon{CKit}~K₁~K₂~K₁⊔K₂}
  describes the operations necessary for definining the composition of a
  map from \ACode{K₁} with a map from \ACode{K₂} that results in a map from
  \ACode{K₁⊔K₂}:
  \LibCode\KCKit
  \noindent The third parameter \ACode{K₁⊔K₂} can be seen as a functional
  dependency\cite{DBLP:conf/esop/Jones00} and is determined
  by the choice of \ACode{K₁} and \ACode{K₂}.
  The fields of a compose kit have the following meaning:
  \begin{itemize}
  \item
    The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
    \ACode{x/t} (according to \ACode{K₁}) and a renaming or substitution
    \ACode{ϕ} (according to \ACode{K₂}) and applies \ACode{ϕ} to \ACode{x/t} resulting in
    a variable or term (according to \ACode{K₁⊔K₂}).
    From this operation we derive map composition \ACode{\ADef{\_·ₖ\_}} as
    shown in the previous subsection.
    % TODO: use the notation def to the first time we use it?
    We use the notation \ACode{ϕ₁ \ADef{·[} C \ADef{]} ϕ₂}, if we need to
    make it explicit that we are using the composition of some
    specific \ACode{\ACon{CKit}} \ACode{C}.
  \item 
    The \ACode{\AField{\&/⋯-⋯}} lemma describes the behavior of
    \ACode{\AField{\_\&/⋯\_}} in terms of \ACode{\AField{\_⋯\_}}, allowing
    subsequent lemmas to make use of the lemmas that we have already proved
    for \ACode{\AField{\_⋯\_}}.
  \item 
    The \ACode{\AField{\&/⋯-wk-↑}} lemma states that applying a map and then
    weakening is the same as weakening first and then lifting the map over
    the variable introduced by the weakening.
    From this lemma, we can derive that lifting distributes over composition:
    \LibCode\KDistLiftCompose
    % \LibCode\KDistLiftComposeProof
  \end{itemize}

  A \ACode{\ACon{CTraversal}} provides a generic \ACode{\AField{fusion}} lemma
  that works for any composition described by a \ACode{\ACon{CKit}}:
  \LibCode\KCTraversal

  Given a \ACode{\ACon{CTraversal}}, we can prove the lemmas about interactions
  of multiple maps, which are required for substitution preserves typing:
  \begin{itemize}
  \item 
    A map \ACode{ϕ} followed by a weakening is equivalent to a
    weakening followed by \ACode{ϕ} that has been lifted over the weakened
    variable:
    \LibCode\KCommLiftWeakenTraverse

    % A weakening followed by a singleton substitution cancel each other out:
    % \LibCode\KWeakenCancelsSingle
    % % \LibCode\KWeakenCancelsSingleProof
    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
    % this lemma, we get the same statement for multiple map applications:
    % \LibCode\KWeakenCancelsSingleTraverse
    % % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton map can be swapped with any map \ACode{ϕ}:
    \LibCode\KDistLiftSingleTraverse
    % A singleton map can be swapped with any map \ACode{ϕ}:
    % \LibCode\KDistLiftSingle
    % % \LibCode\KDistLiftSingleProof
    % By applying \ACode{\AField{fusion}} to this lemma, we get the
    % same statement for multiple map applications:
    % \LibCode\KDistLiftSingleTraverse
    % % \LibCode\KDistLiftSingleTraverseProof
  \end{itemize}
  We define \ACode{\ACon{CKit}} instances for each of the four
  combinations between renamings and substitutions:
  \LibCode\KCKitInstances
  Interestingly, the \ACode{\ADef{Cₛ}} instance has another
  \ACode{\ADef{CKit}} parameter.
  This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
  which yields the \ACode{\ADef{CKit}} between substitutions
  and renamings, and then use this resulting \ACode{\ADef{CKit}}
  as an argument for the case with two substitutions.
  As shown below, Agda's instance resolution is up for the task:

  % To get the \ACode{\ADef{CKit}} between substitutions and
  % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
  % The resulting \ACode{\ADef{CKit}} is then again used as an argument
  % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
  % between two substitutions. As shown below, Agda's instance
  % resolution is up for the task:
  \noindent
  \begin{minipage}[t]{0.5\linewidth}
  \LibCode\KCKitInstancesConcreteI
  \end{minipage}
  \begin{minipage}[t]{0.5\linewidth}
  \LibCode\KCKitInstancesConcreteII
  \end{minipage}

  % \LibCode\KCKitNotation

@@

  For this purpose, the \ACode{\ACon{CKit}} abstraction, needs
  to satisfy two requirements:
  it needs to provide us with enough machinery to define composition, but
  it also needs to provide us with enough axioms that we are able to instantiate
  the \ACode{\ACon{CTraversal}}. It is defined as followed:
  \LibCode\KCKit
  The \ACode{\AField{\_\&/⋯\_}} operation takes a variable or term
  (according to \ACode{K₁}) and a renaming or substitution
  (according to \ACode{K₂}) and performs the map traversal returning
  a variable or term (according to \ACode{K₁⊔K₂}).
  The third parameter \ACode{K₁⊔K₂} can be seen as a functional
  dependency\cite{DBLP:conf/esop/Jones00} and is determined
  by the choice of \ACode{K₁} and \ACode{K₂}.
